---
title: "Assignment-1"
format: html
---

This page contains my Assignment 1 work.

---

## Video Summary

The video discusses the relationship between Data Science and Industrial Engineering. 
The speakers explain how data science techniques can be used to improve decision-making in engineering systems and business processes. 

They emphasize that modern organizations generate large amounts of data and industrial engineers can use this data to optimize production systems, logistics, and operations. 
The talk also highlights the importance of programming, statistics, and analytical thinking for engineers who want to work in data-driven environments.

Overall, the video shows how data science tools help industrial engineers analyze complex systems and make more efficient and informed decisions.

---

## Open-ended Question

How can data science improve decision-making in industrial engineering?

**Answer**

Data science helps industrial engineers analyze large datasets, identify patterns, and optimize processes. By using statistical analysis and machine learning techniques, engineers can make better predictions and improve operational efficiency.

---

## Multiple Choice Question

Which skill is essential for combining data science with industrial engineering?

A) Artistic drawing  
B) Statistical analysis  
C) Musical composition  
D) Architecture design  

**Correct Answer:** B) Statistical analysis

---

```{r}
#| label: task-3-c
#| echo: true

my_first <- "Özge"
my_birth_year <- 1996

# k sayısı hesaplama
k <- (nchar(my_first) + my_birth_year) %% 15 + 8
cat("calculated my k :", k, "\n")

# function
compute_stats <- function(x) {
  list(
    mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    variance = var(x, na.rm = TRUE),
    iqr = IQR(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE)
  )
}

# 1. using sapply
cat("\n--- sapply Sonuçları ---\n")
sapply(mtcars, compute_stats)

# 2. using apply
cat("\n--- apply Sonuçları ---\n")
apply_results <- apply(as.matrix(mtcars), 2, compute_stats)
print(apply_results)

# for döngüsü ile mtcars analizi 
cat("\n--- For Döngüsü Sonuçları ---\n")
for (i in 1:ncol(mtcars)) {
  cat(colnames(mtcars)[i], ":\n")
  print(compute_stats(mtcars[[i]]))
}
```

```{r}
#| label: task-3-c
#| echo: true

# Personal variables
my_first <- "Feyza"
my_birth_year <- 1990

# Calculate k
k <- (nchar(my_first) + my_birth_year) %% 15 + 8
cat("Calculated value of k:", k, "\n")

# Custom statistical summary function
compute_stats <- function(x) {
  list(
    mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    variance = var(x, na.rm = TRUE),
    IQR = IQR(x, na.rm = TRUE),
    minimum = min(x, na.rm = TRUE),
    maximum = max(x, na.rm = TRUE)
  )
}

# Using a for loop
cat("\n--- Results using a for loop ---\n")

for (i in 1:ncol(mtcars)) {
  cat("\nColumn:", colnames(mtcars)[i], "\n")
  print(compute_stats(mtcars[[i]]))
}

# Using sapply
cat("\n--- Results using sapply ---\n")
sapply(mtcars, compute_stats)

# Using apply
cat("\n--- Results using apply ---\n")
apply(as.matrix(mtcars), 2, compute_stats)
 ⁠

⁠ {r}
#| label: task-3-d
#| echo: true

library(dslabs)
data("polls_us_election_2016")

cat("\nValue of k:", k, "\n")

# Display k rows of the original dataset
cat("\n--- k rows from the original dataset ---\n")

if (k %% 2 == 0) {
  print(head(polls_us_election_2016, k))
} else {
  print(tail(polls_us_election_2016, k))
}

# Total number of NA values
cat("\nTotal number of NA values in the dataset:\n")
print(sum(is.na(polls_us_election_2016)))

# NA values per column (sorted)
na_per_column <- sort(colSums(is.na(polls_us_election_2016)), decreasing = TRUE)

cat("\nNA values per column (sorted largest to smallest):\n")
print(na_per_column)

# Top 8 columns with most NA values
cat("\nTop 8 columns with the highest number of NA values:\n")
print(head(na_per_column, 8))

# Create new dataset without overwriting original
new_data <- polls_us_election_2016

# Replace NA values
for (i in 1:ncol(new_data)) {
  
  if (is.numeric(new_data[[i]])) {
    
    new_data[[i]][is.na(new_data[[i]])] <- my_birth_year + k
    
  } else if (is.character(new_data[[i]])) {
    
    new_data[[i]][is.na(new_data[[i]])] <- paste0(my_first, "_", k)
    
  } else if (is.factor(new_data[[i]])) {
    
    temp <- as.character(new_data[[i]])
    temp[is.na(temp)] <- paste0(my_first, "_", k)
    new_data[[i]] <- factor(temp)
    
  }
}

# Display k rows of the new dataset
cat("\n--- k rows from the new dataset ---\n")

if (k %% 2 == 0) {
  print(head(new_data, k))
} else {
  print(tail(new_data, k))
}

# Remaining NA values
cat("\nTotal NA values remaining in new_data:\n")
print(sum(is.na(new_data)))

# Check if any NA remains
cat("\nDoes new_data still contain NA values?\n")
print(anyNA(new_data))
 ⁠

	⁠*AI Disclosure:* I utilized AI tools (Gemini/ChatGPT) to assist in structuring the R code and translating the assignment requirements.
>
	⁠*Prompt used:* "Could you help me write R code for my EMU660 assignment that calculates k, analyzes the mtcars dataset, computes summary statistics, and replaces NA values in the polls_us_election_2016 dataset?"
