---
title: "Assignment-1"
format: html
---

This page contains my Assignment 1 work.

---

## Video Summary

The video discusses the relationship between Data Science and Industrial Engineering. 
The speakers explain how data science techniques can be used to improve decision-making in engineering systems and business processes. 

They emphasize that modern organizations generate large amounts of data and industrial engineers can use this data to optimize production systems, logistics, and operations. 
The talk also highlights the importance of programming, statistics, and analytical thinking for engineers who want to work in data-driven environments.

Overall, the video shows how data science tools help industrial engineers analyze complex systems and make more efficient and informed decisions.

---

## Open-ended Question

How can data science improve decision-making in industrial engineering?

**Answer**

Data science helps industrial engineers analyze large datasets, identify patterns, and optimize processes. By using statistical analysis and machine learning techniques, engineers can make better predictions and improve operational efficiency.

---

## Multiple Choice Question

Which skill is essential for combining data science with industrial engineering?

A) Artistic drawing  
B) Statistical analysis  
C) Musical composition  
D) Architecture design  

**Correct Answer:** B) Statistical analysis

---
```{r}
#| label: task-3-c
#| echo: true

# Personal variables
my_first <- "Feyza"
my_birth_year <- 1990

# Calculate k
k <- (nchar(my_first) + my_birth_year) %% 15 + 8
cat("Calculated value of k:", k, "\n")

# Custom statistical summary function
compute_stats <- function(x) {
  list(
    mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    variance = var(x, na.rm = TRUE),
    iqr = IQR(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE)
  )
}

# Using sapply
cat("\n--- sapply results ---\n")
sapply(mtcars, compute_stats)

# Using apply
cat("\n--- apply results ---\n")
apply_results <- apply(as.matrix(mtcars), 2, compute_stats)
print(apply_results)

# Using for loop
cat("\n--- For loop results ---\n")
for (i in 1:ncol(mtcars)) {
  cat(colnames(mtcars)[i], ":\n")
  print(compute_stats(mtcars[[i]]))
}

### Task 3-d
```{r}
#| label: task-3-d
#| echo: true

library(dslabs)
data("polls_us_election_2016")

data_original <- polls_us_election_2016

# Display k rows
cat("\nRows from original dataset:\n")

if (k %% 2 == 0) {
  print(head(data_original, k))
} else {
  print(tail(data_original, k))
}

# Total NA count
cat("\nTotal NA values:", sum(is.na(data_original)), "\n")

# NA per column
na_per_column <- sort(colSums(is.na(data_original)), decreasing = TRUE)

cat("\nNA values per column (top 8):\n")
print(head(na_per_column, 8))

# Replace NA values
new_data <- data_original

for(i in 1:ncol(new_data)) {

  if(is.numeric(new_data[[i]])) {

    new_data[[i]][is.na(new_data[[i]])] <- my_birth_year + k

  } else {

    new_data[[i]] <- as.character(new_data[[i]])

    new_data[[i]][is.na(new_data[[i]])] <- paste0(my_first, "_", k)

  }

}

# Display rows from new dataset
cat("\nRows from new dataset:\n")

if (k %% 2 == 0) {
  print(head(new_data, k))
} else {
  print(tail(new_data, k))
}

# Remaining NA
cat("\nRemaining NA values:", sum(is.na(new_data)), "\n")

# Check any NA
cat("\nAny NA left:", anyNA(new_data), "\n")

```

**AI Disclosure**

I used AI tools (ChatGPT) to help structure the R code and better understand the assignment requirements.

Prompt used:  
"Write R code for an assignment that calculates k, analyzes mtcars dataset with cus
